---
title: "Bayesian_algorithm"
author: "Pedro Valls"
date: '2022-11-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Some Bayesian calculations

We are going to give some examples of the Bayes' theorem using simulated data.
Remember that Bayes' Theorem is given by 
$ p(\theta \mid y) = \frac{\ell(y \mid \theta) \cdot p(\theta)}{p(y)}$
where 
$p(\theta \mid y)$ is the posterior distribution,
$\ell(y \mid \theta)$ is the likelihood, and 
$p(y)$ is the marginal distribution.

# Simulating data for a regression

First define the sample size $n$, the $\beta$ coefficient and the $\tau$ the precision which is the inverse of the variance $\sigma^{2}$. 

```{r}
n <- 50  
beta <- 0.9 
tau <- 1.0  
```

Now define the regressors $y=(y_{1}, \cdots, y_{n})$ as $U[10,20]$, i.e. draw $n$ uniform random number in the interval $[10,20]$.

```{r}
set.seed(123456)
y<- runif(n,min=10,max=20)
print(y)
```

Now define the dependent variable as drawn from a normal random variable with mean $\beta \cdot y$ and variance equal to $\frac{1}{\tau}$.

```{r}
set.seed(123456)
consump <- rnorm(n,mean = beta*y,sd =1/sqrt(tau))
print(consump)
```

# Plotting the likelihood
Since the likelihood is $\ell(y \mid \beta)$ we need a range of possible values for $\beta$ that are close to the true value. 
Let us use $\beta=[0.86, \cdots, 0.94]$ using $100$ numbers in this interval.
Let us compute $\hat{\beta}$ the OLS estimate if $\beta$. 
Plot the scatter plot of y $\times$ consump
Plot the likelihood
```{r}
par(mfrow=c(1,2))
betah <- sum(consump*y)/sum(y*y)
betavalues <- seq (0.86, 0.94, length = 100)
require(stats)
z <- lm(consump ~ y - 1)
plot(y, consump)
abline(z,col="red")
plot(betavalues, dnorm(betavalues, betah, 1/sqrt(tau*sum(y*y))),type ="l", ylab="likelihood", xlab = "beta")
```

